A Recipe for Scavenging - the natural history of a behaviour

### 2016-04-01 natural history of scavenging.tex
seabird details

\affiliations\{
\item A. Kane (\url{adam.kane@ucc.ie}), University College Cork, Cooperage Building, School of Biological Earth and Environmental Sciences, Cork, Ireland.
\item K. Healy and A. Jackson, Trinity College Dublin, Department of Zoology, School of Natural Sciences, Dublin Ireland.
\item T. Guillerme, Imperial College London, Silwood Park Campus, Department of Life Sciences, Buckhurst Road, Ascot SL5 7PY, United Kingdom.
\item G. Ruxton, School of Biology, Sir Harold Mitchell Building, Greenside Place, St Andrews, KY16 9TH, United Kingdom.
}

@@

\maketitle

@@

Of course, tetrapod terrestrial dominance predates the evolution of the dinosaurs. %TG: I will down that "dominance" term. Maybe change the sentence to "Of course the importance of terrestrial tetrapods in ecosystems predates the evolution of the dinosaurs."
It is during the Permian (298.9 - 252.17 Mya) that we have the earliest evidence of vertebrate scavenging where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}. 
And it is with the evolution of endothermy in the therapsid-mammal lineage \citep{clarke2010temperature} that terrestrial vertebrates would have gained the ability to range widely, a vital component in seeking out carrion.

@@

%TG: I think it's missing a transition here: maybe something along "But moving on land as it's limitations..."
But it is in the air that we find scavengers \textit{par excellence}. %TG: not sure how reviewers will go with French but I like it!
Flight is a cheaper means of locomotion than either walking or running \citep{tucker1975energetic}. 
We know that many extant birds exist as facultative scavengers because storks, raptors and corvids all take substantial quantities of carrion in their diet \citep{kendall2013alternative}.
The advantage of flight can be extended further in larger species that engage in soaring instead of flapping flight, which is even cheaper still \citep[pproximately twice the basal metabolic rate; ][]{hedenstrom1993migration,spivey2014analysing}.
The benefits this confers are clear from the information we have on the enormous foraging ranges of many vultures \citep{spiegel2013factors} and seabirds \citep{thaxter2012seabird}.
In vultures we have the best known scavengers on Earth. 
These birds consist of two convergent groups, from the old and the new world where they represent the only example of obligate vertebrate scavengers today. %TG: Or "obligate vertebrate scavengers known to date." For a more emphasizing turn.

@@

The only other vertebrate group capable of powered flight are the bats where scavenging has not been recorded to our knowledge.
However, some fossils evidences as early as in the Eocene (56 to 33.9 Mya) suggest the presence of possible scavenging bats.
A big bat from the middle to late Eocene (56 to 33.9 Mya) bat with a robust cranio-mandibular morphology, \textit{Necromatis} (``death-eater''), could have intake carrion as a decent proportion of its diet \citep{Weithofer_Necromantis_1887,Hand_Necromantis_2012}.
However, this example might be peculiar to this exceptional fossil since most other fossils seems less suited to this lifestyle \citep{eiting2009global}. 
%Although it does not seem that flight is the main criterion precluding them from scavenging (see below).
%TG: Actually, maybe this example fits better in the food processing part (and maybe change the silhouettes in the picture then)?
%``The name Necromantis (``death-eater'') alludes to the diet that Weithofer assumed for this bat – carrion, but probably also small, live vertebrates, including mammals such as rodents and other bats.'' from the paper \cite{Hand_Necromantis_2012}.

@@

Large body size confers substantial dominance and starvation-resistance benefits \citep{ruxton2004obligate}.
Thus, we would expect scavengers to have this trait selected for even in the case of weight-constrained fliers.
Cinereous Vultures (\textit{Aegypius monachus}) and condors (\textit{Vultur gryphus}, \textit{Gymnogyps californianus}) all have body masses that can exceed 10 kg and represent some of the heaviest bird species capable of flight \citep{ferguson2001raptors,donazar2002effects}.

And as we have noted the Azhdarchid pterosaurs were far bigger again, with estimated body masses of over 200 kg \citep{witton2010size}. Although \cite{witton2008reappraisal} argued that neck inflexibility and straight, rather than hooked jaw morphology points against pterosaurs existing as \textit{obligate} scavengers, Azhdarchid terrestrial proficency indicates they would have been comfortable foraging on the ground.
Indeed, extant Marabou Storks (\textit{Leptoptilos crumenifer}) %TG: Check if that's the one
have a comparable morphology and are noted facultative scavengers so it is reasonable to believe that certain pterosaurs behaved similarly.

>>>>

%\affiliations\{
%\item A. Kane (\url{adam.kane@ucc.ie}), University College Cork, Cooperage Building, School of Biological Earth and Environmental Sciences, Cork, Ireland.
%\item K. Healy and A. Jackson, Trinity College Dublin, Department of Zoology, School of Natural Sciences, Dublin Ireland.
%\item T. Guillerme, Imperial College London, Silwood Park Campus, Department of Life Sciences, Buckhurst Road, Ascot SL5 7PY, United Kingdom.
%\item G. Ruxton, School of Biology, Sir Harold Mitchell Building, Greenside Place, St Andrews, KY16 9TH, United Kingdom.
%}

@@

% \maketitle

@@

Of course, the importance of terrestrial tetrapods predates the evolution of the dinosaurs. %TG: I will down that "dominance" term. Maybe change the sentence to "Of course the importance of terrestrial tetrapods in ecosystems predates the evolution of the dinosaurs." AK Done 
It is during the Permian (298.9 - 252.17 Mya) that we have the earliest evidence of vertebrate scavenging where a temnospondyl amphibian fed on the carcass of \textit{Varanops}, a predatory synapsid of the time \citep{reisz2006articulated}. 
And it is with the evolution of endothermy in the therapsid-mammal lineage \citep{clarke2010temperature} that terrestrial vertebrates would have gained the ability to range widely, a vital component in seeking out carrion.

@@

%TG: I think it's missing a transition here: maybe something along "But moving on land as it's limitations..."
But it is in the air that we find scavengers \textit{par excellence}. %TG: not sure how reviewers will go with French but I like it!
Flight is a cheaper means of locomotion than either walking or running \citep{tucker1975energetic}. 
We know that many extant birds exist as facultative scavengers because storks, raptors and corvids all take substantial quantities of carrion in their diet \citep{kendall2013alternative}.
The advantage of flight can be extended further in larger species that engage in soaring instead of flapping flight, which is even cheaper still \citep[pproximately twice the basal metabolic rate; ][]{hedenstrom1993migration,spivey2014analysing}.
The benefits this confers are clear from the information we have on the enormous foraging ranges of many vultures \citep{spiegel2013factors} and seabirds \citep{thaxter2012seabird}.
The difficulty inherent in observing marine birds foraging at sea means our knowledge of their feeding habits is limited. 
But it does seem that albatrosses, who can range many 100s of kilometres, take a substantial amount of carrion in their diet \citep{croxall1994dead}.
This is typically in the form of squid, which float on the surface, allowing the birds to readily pluck their remains out of the water \citep{croxall1994dead}. 
In vultures, who are more easily observed on land, we have the best known scavengers on Earth. 
These birds consist of two convergent groups, from the old and the new world where they represent the only example of obligate vertebrate scavengers to date. %TG: Or "obligate vertebrate scavengers known to date." For a more emphasizing turn.

@@

The only other vertebrate group capable of powered flight are the bats where scavenging has not been recorded to our knowledge.
However, some fossils evidence as early as in the Eocene (56 to 33.9 Mya) suggest the presence of possible scavenging bats.
A large bat from the middle to late Eocene (56 to 33.9 Mya) with a robust cranio-mandibular morphology, \textit{Necromatis} (``death-eater''), could have intake carrion as a significant proportion of its diet \citep{Weithofer_Necromantis_1887,Hand_Necromantis_2012}.
However, this example might be peculiar to this exceptional fossil given most other fossils represent animals less suited to the lifestyle \citep{eiting2009global}.

@@

Large body size confers substantial dominance and starvation-resistance benefits \citep{ruxton2004obligate}.
Thus, we would expect scavengers to have this trait selected for even in the case of weight-constrained fliers.
Wandering Albatrosses (\textit{Diomedea exulans}), Cinereous Vultures (\textit{Aegypius monachus}) and condors (\textit{Vultur gryphus}, \textit{Gymnogyps californianus}) all have body masses that can exceed 10 kg and represent some of the heaviest bird species capable of flight \citep{weimerskirch1992reproductive,ferguson2001raptors,donazar2002effects}.

And as we have noted the Azhdarchid pterosaurs were far bigger again, with estimated body masses of over 200 kg \citep{witton2010size}. Although \cite{witton2008reappraisal} argued that neck inflexibility and straight, rather than hooked jaw morphology points against pterosaurs existing as \textit{obligate} scavengers, Azhdarchid terrestrial proficency indicates they would have been comfortable foraging on the ground.
Indeed, extant Marabou Storks (\textit{Leptoptilos crumenifer}) %TG: Check if that's the one
have a comparable morphology and are noted facultative scavengers \citep{monadjem2012survival} so it is reasonable to believe that certain pterosaurs behaved similarly.
